DATA 621---Business Analytics and Data Mining

### 2020-09-21 HW1/HW1_Group2.Rmd
More copyediting

```{r model2C+S}
##### Model 2: Pairwise Model
# cleaning data for model 2
ts2 <- train_set[, -(17:19)]
ts2 <- data.frame(apply(ts2, 2, scale))
ts2$TARGET_WINS <- train_set$TARGET_WINS
```

@@

```{r model3}

##### Model 3: K-fold Model
stepdata <- subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
stepdata <- cbind(stepdata, sapply(stepdata[2:length(stepdata)], function(x) x^2))
names(stepdata) <- make.unique(names(stepdata), sep = "_")
model <- lm(TARGET_WINS ~ ., data = stepdata)
cooksd <- cooks.distance(model)
influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
stepdata.2 <- stepdata[-c(influential), ]
# Set up repeated k-fold cross-validation
set.seed(525)
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model <- train(TARGET_WINS ~ ., data = stepdata.2, method = "lmStepAIC", 
                   trControl = train.control, trace = FALSE)
# Final model 
step.model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BASERUN_SB + 
                  TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_FIELDING_E + 
                  TEAM_FIELDING_DP + TEAM_BATTING_BB_1 + TEAM_BATTING_SO_1 + 
                  TEAM_BASERUN_CS_1 + TEAM_PITCHING_H_1 + TEAM_PITCHING_BB_1 + 
                  TEAM_FIELDING_E_1 + TEAM_FIELDING_DP_1 + TEAM_BATTING_1B_1, 
                data = stepdata.2)
# Summary of the model
mod_sum3 <- summary(step.model)
```

@@

From studying each coefficient, this discussion begins with the variables that follow 
the logic of winning and are self-evident. Firstly, the model suggests that winning is 
in favor of the team batting hits more bases, (1B, 2B, 3B, HR). It further reveals that 
an increase in singles would lead to better chances of winning. A hit is a hit, and more 
hits than the opponent will inevitably lead to a win. Likewise, an increase in walks by 
the batting team is awarded first base, further increasing base counts and, thus the 
chances of winning. In terms of the base run, an increase in the number of stolen bases, 
and a decrease in caught steals would lead to a win for the batting team. These variables 
have a positive effect on the number of wins, and it is understandable as to why.

>>>>

```{r model2C+S}
# Model 2: Pairwise Model
# Cleaning data for model 2
ts2 <- train_set[, -(17:19)]
ts2 <- data.frame(apply(ts2, 2, scale))
ts2$TARGET_WINS <- train_set$TARGET_WINS
```

@@

```{r model3}
# Model 3: K-fold Model
stepdata <- subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
stepdata <- cbind(stepdata,
                  sapply(stepdata[2:length(stepdata)], function(x) x^2))
names(stepdata) <- make.unique(names(stepdata), sep = "_")
model <- lm(TARGET_WINS ~ ., data = stepdata)
cooksd <- cooks.distance(model)
influential <- as.numeric(
  names(cooksd)[(cooksd > 4 * mean(cooksd, na.rm = TRUE))])
stepdata.2 <- stepdata[-c(influential), ]

# Set up repeated k-fold cross-validation
set.seed(525)
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model <- train(TARGET_WINS ~ ., data = stepdata.2, method = "lmStepAIC", 
                   trControl = train.control, trace = FALSE)
# Final model 
step.model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B +
                   TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H +
                   TEAM_PITCHING_HR + TEAM_FIELDING_E +  TEAM_FIELDING_DP +
                   TEAM_BATTING_BB_1 + TEAM_BATTING_SO_1 +  TEAM_BASERUN_CS_1 +
                   TEAM_PITCHING_H_1 + TEAM_PITCHING_BB_1 +  TEAM_FIELDING_E_1 +
                   TEAM_FIELDING_DP_1 + TEAM_BATTING_1B_1, data = stepdata.2)

# Summary of the model
mod_sum3 <- summary(step.model)
```

@@

This discussion begins with the variables that follow the logic of winning and
are self-evident. Firstly, the model suggests that winning is in favor of the
team which when batting has more total bases, (1B, 2B, 3B, HR). It further
reveals that an increase in singles would lead to better chances of winning. A
hit is a hit, and more hits than the opponent will inevitably lead to a win.
Likewise, an increase in walks by the batting team is awarded first base,
further increasing base counts and, thus the chances of winning. In terms of the
base run, an increase in the number of stolen bases, and a decrease in caught
steals would lead to a win for the batting team. These variables have a positive
effect on the number of wins, and it is understandable as to why.
